---
output: 
  pdf_document: 
    fig_caption: true
    fig_width: 6
    fig_height: 4

params:
  date: !r Sys.Date()
  # target.label: "PERF.all"
  target.label: "PERF09"
  features.set: "big5items"
  # features.set: "big5composites"
  split.ratio: 0.80
  cv.repeats: 100

title: "Job Performance Analysis with
        target = `r params$target.label` and features set = `r params$features.set`"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# knitr::opts_knit$set(global.par = TRUE)
options(digits = 3)

# clear the workspace
# rm(list=ls()) # tricky: deletes params

dataset.label <- paste0(c("data/dataset", "rds"), collapse = ".")

# devtools::install_github("agilebean/machinelearningtools", force = TRUE)
# load libraries
libraries <- c("dplyr", "magrittr", "tidyverse"
               , "sjlabelled" # read SPSS
               , "caret", "doParallel"
               , "DataExplorer", "RColorBrewer"
               , "machinelearningtools"
               , "knitr", "pander"
)
sapply(libraries, require, character.only = TRUE)

# nominal <- FALSE # with ordinal as ORDERED factors
nominal <- TRUE # with ordinal as NOMINAL factor

seed <- 17

get_features <- function(target_label, features_set, data_set) {
  
  data_set %>% 
    select(-target_label,
           -starts_with("TO"),
           -starts_with("PERF")
    ) %>% 
    {
      if (features_set == "big5items") {
        # remove composite scores - equivalent to (-nn, -ee, -oo, -aa, -cc)
        select(., -matches("(oo|cc|ee|aa|nn)$")) 
        
      } else if (features_set == "big5composites") {
        # remove Big5 items
        select(., -matches(".*(1|2|3|4|5|6)"))
        
      } else { . }
    } %>% 
    names %T>% print
}
```


# 1. Data Acquistion

```{r data acquisition, include=TRUE}
#######################################################################
# 1. Data Acquistion
#######################################################################
dataset.label %>% print
dataset <- readRDS(dataset.label)

```

# 2. Data Preparation

## 2.4 Select the target & features

```{r select target & features code, include=FALSE}
########################################
## 2.4 Select the target & features
########################################

set.seed(seed)
dataset %<>% nrow %>% sample %>% dataset[.,]

########################################
# 3.2: Select the features & formula
########################################
# define features
features <- get_features(params$target.label, params$features.set, dataset)

# define formula
formula1 <- set_formula(params$target.label, features)

# grepl("PERF[0-9]", params$target.label) %>% print
# r  {paste("the averaged performance score")}
# if (params$target.label == "PERF.all") # good
# ìf (params$target.label == "PERF.all") #  bad

```

For this analysis, the target is set to `r if (params$target.label == "PERF.all") {paste("the averaged performance score")} else if (grepl("PERF[0-9]", params$target.label)) { paste("a one-year performance score")} else if (params$target.label == "TO.all") {paste("the averaged turnover score")} else {paste(params$target.label)}`, and the features are `r if (params$features.set == "big5items") {paste("all Big5 items")} else if (params$features.set =="big5composites") {paste("the Big5 composite scores")} else {paste(params$features.set)}`.

The number of variables is reduced from `r nrow(dataset)`  variables to the target variable (`r params$target.label`) and `r length(features)` features.

```{r select target & features, echo=TRUE}
params$target.label
features
```

## 2.3 Split the data
To train the machine learning models, we split the data into a training set (`r params$split.ratio*100`%) for training and tuning the models, and a testing set (`r 100-params$split.ratio*100`%) that serves as unseen data for the final validation step.
To avoid imbalance between training and testing set, we must first randomize the data (shuffle).


```{r split the data code, include=FALSE}
########################################
## 2.3 Split the data
########################################
# shuffle data
set.seed(seed)
# set.seed(5)
shuffle.index <- dataset %>% nrow %>% sample
dataset %<>% .[shuffle.index,] %T>% print
# short version:
# dataset %<>% nrow %>% sample %>% dataset[.,] %T>% print

# later: imputation of NAs
# dataset %>% preProcess(method="knnImpute") %>% print

# dataset subsetting for tibble: [[
training.index <- createDataPartition(dataset[[params$target.label]], 
                                      p = params$split.ratio, list = FALSE)
testing.set <- dataset[-training.index, ]
training.set <- dataset[training.index, ]

```

We verify the `r params$split.ratio*100`:`r 100-params$split.ratio*100` split by checking the number of observations:

```{r split the data}
training.set %>% nrow
testing.set %>% nrow
```

# 3. Train Model

```{r train the model code, include=FALSE}
################################################################################
# 3. Train Model
# 3-1: Select a model
# 3-2: Select the target, features, training data
# 3-3: Train the model with the target and features
################################################################################

########################################
# 3-1: Select a model
########################################
algorithm.list <- c(
  "lm",
  "knn",
  "gbm",
  "rf",
  "ranger",
  "xgbTree",
  "xgbLinear",
  "svmLinear",
  "svmRadial"
)
  
```

The machine learning models used in the benchmarking are:

**Linear models:**
  | "lm": linear regression (benchmarking baseline)
  | "glm": logistic regression
  | "xgbLinear": extreme gradient boosting  - linear kernel
  | "svmLinear": support vector machines - linear kernel
  
**Tree-based models:**
  | "gbm": gradient boosting machines
  | "rf", "ranger": random forests
  | "xgbTree": extreme gradient boosting - tree kernel
  
**Other non-linear models:**
  | "knn": k nearest neighbors
  | "svmRadial": support vector machines - radial kernel

# 4. Evaluate Models
```{r get model, cache=TRUE, include=FALSE}
models.list.name <- paste("data/models.list",
                          params$target.label, 
                          params$features.set,
                          paste0(params$cv.repeats, "repeats"),
                          "rds",
                          sep = ".") %T>% print
models.list <- models.list.name %>% readRDS

training_configuration_number <- models.list[[1]] %>% .$control %>% .$number
training_configuration_repeats <- models.list[[1]] %>% .$control %>% .$repeats
```

Training the model on the training set with `r training_configuration_number`-fold cross-validation, repeated  `r training_configuration_repeats` times, yields the following results. Linear regression as benchmark reference is denoted as "lm".

## 4.1 Training Set Performance

```{r train the model,fig.width=7, fig.height=5, cache=FALSE, echo=FALSE}
########################################
## 4.1 Training Set Performance
########################################
# get model metrics
models.metrics <- models.list %>% get_model_metrics 

# display Rsquared trainingset performance - table
# models.metrics$Rsquared.training %>% kable(caption = "training set performance: Rsquared")

# display RMSE trainingset performance - table
models.metrics$RMSE.training %>% kable(caption = "training set performance: RMSE")

# display RMSE trainingset performance - boxplots
models.metrics$RMSE.boxplots

# models.metrics$RMSE.training %>% filter(str_detect(model, "^lm$"))
rank.lm.training <- which(models.metrics$RMSE.training$model == "lm")
```

We can see that linear regression ("lm") performs on `r rank.lm.training`th RMSE rank of the selected machine learning models. This result must be validated on the testing set.


## 4.2 Testing Set Performance

```{r testing set RMSE, cache=FALSE}
########################################
## 4.2 Testing Set Performance
########################################

## very tricky: models.list is overridden by previous code
## solution: read again
# models.list <- models.list.name %>% readRDS
# 
# # RMSE for all models on testing set
# models.list %>%
#   get_model_metrics %>%
#   .$RMSE.testing %>% kable(caption = "testing set performance: RMSE")
# 
# # training vs. testing set performance
# models.list %>%
#   get_model_metrics %>%
#   .$RMSE.all %>% kable(caption = "training vs. testing set performance: RMSE")

models.metrics$RMSE.testing %>% kable(caption = "testing set performance: RMSE")

rank.lm.testing <- which(models.metrics$RMSE.testing$model == "lm")
```

We had reserved a portion of the original data as testing set to validate the model performance on unseen data. 
The model comparison on RMSE shows that linear regression model `r if (rank.lm.testing == rank.lm.training) {paste0("remained on ", rank.lm.training, "th rank.")} else if (rank.lm.testing < rank.lm.training) {paste0("improved from ", rank.lm.training, "th to ", rank.lm.testing, "th rank.")} else {paste0("dropped from ", rank.lm.training, "th to ", rank.lm.testing, "th rank.")}`

The following table compares all models training set performance with their testing set performance.

```{r testing set benchmark all}
models.metrics$RMSE.all %>% kable(caption = "training vs. testing set performance: RMSE")
```









